---
title: "Stats506_PS3"
format:
  html:
    embed-resources: true
editor: visual
---

# Problem 1 - Vision

### Part A:

I will download the VIX_D file and DEMO_D file. I will merge the two files to create a single Stata dataset, using the SEQN variable for merging. I will keep only records which match. Below I will print my total size, showing that it is now 6,980.

### Part B:

Without fitting any models, I will estimate the proportion of respondents within each 10-year age bracket (eg. 0-9, 10-19, 20-29, etc) who wear glasses/contact lenses for distance vision. I will produce a nice table with the results below. (There was hint)

### Part C:

Fit three logistic regression models predicting whether a respondent wears glasses/contact lenses for distance vision. Predictors:

1.  age
2.  age, race, gender
3.  age, race, gender, poverty income ratio

Next, I will produce a table presenting the estimated odds ratios for the coefficients in each model along with the sample size for the model, the pseud-R\^2, and AIC values.

### Part D:

From the third model from the previous part, I will discuss whether the *odds* of men and women being wears of glasses/contact lenses for distance vision differ. I will test whether the *proportion* of wearers of glasses/contact lenses for distance vision differs between men and women. I will include the results of the test and its interpretation.

# Problem 2 - Sakila

Load the "sakila" database discussed in class into SQLite. It can be downloaded from the link provided.

```{r}
library(DBI)
library(RSQLite)
```

```{r}
#Load the "sakila" database into SQLite
sakila <- dbConnect(RSQLite::SQLite(), "/Users/19892/OneDrive/Documents/STATS506/ProblemSets/Stats506_PS3/sakila_master.db")
```

### Part A:

Aside from English, what language is most common for films? Answer this with a a single SQL query.

To get a feel for the data base, here are the table names

```{r}
dbListTables(sakila)
```

Below are the languages:

```{r}
dbListFields(sakila, "language")
dbGetQuery(sakila, "SELECT name FROM language")
dbGetQuery(sakila, "SELECT language_id from language")
```

```{r}
dbGetQuery(sakila,"
  SELECT language_id, COUNT(*) AS language_count
    FROM film
    GROUP BY language_id
    ORDER BY language_count DESC
           "
           )
```

Above it seems that every film has a language id of 1 which corresponds to English. Therefore, every film was done in English and English is the most common language.

*For each of the following questions, solve them in two ways: First, use SQL query or queries to extract the appropriate table(s), then use regular R to answer the question. Second, use a single SQL query to answer the question.*

### Part B: What genre of movie is the most common in the data, and how many movies are of this genre?

SQL:

```{r}
dbListFields(sakila, "film_category")
dbGetQuery(sakila, "SELECT name FROM category")
```

```{r}
dbGetQuery(sakila,"
  SELECT category_id, COUNT(*) AS genre_count
    FROM film_category
    GROUP BY category_id
    ORDER BY genre_count DESC
           "
           )
```

\*How do I do the next step- category id with genre name?
